Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-02-17 paper/structureScene.tex
Edit results section

%The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation. Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering and MFCCs features.

This section presents evaluation results for the acoustic scene similarity retrieval task. The $p@k$ for different settings are shown in Figure \ref{fig:ASS_1}, illustrating the effect of the different similarity metrics. This allows us to also evaluate the recall capabilities of the system under evaluation.

@@

As can be seen on Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when considering MFCCs as features. Its performance decreases compared to the early approach when considering the scattering features. Extensive analysis of those results is out of the scope of this paper, but we can assume that this decrease is to to the dimensionality reduction operated on the scattering features before GMM training in order to ensure the convergence of the EM algorithm.

The early approach being simpler in terms of implementation and runtime complexity, this method is retained as baseline for the remaining.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/baselines}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. }}
\label{fig:ASS_-1}
\end{center}
\end{figure}

@@

As can be seen on Figure \ref{fig:ASS_0}, the logarithmic compression is beneficial for precision performance irrespective of the similarity considered. The gain is more pronounced when considering alternative similarities.

@@

For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes. However, it is worth noticing that \emph{RbQ-a} performs worse or comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids. This result is in line with the findings of \cite{lagrange:hal-01082501}. To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects. Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes compared to \emph{RbQ-w}. This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution. Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames. On the contrary, two similar environments may share a lot of similar sound sources with only a few sources discriminating between them. \ml{This is a bit of a stretch, since it may just be that we used the wrong distance to take the weighting into account. Perhaps the non-normalized EMD is not appropriate here?}

\ja{needs to be discussed}

Considering $p@5$ as our metric (as in \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), the use of the log-scattering transform instead of MFCCs increases the performance from $0.31$ to $0.49$ and the use of the relevance-based quantization approach using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$. \vl{Too many instances of ``use'', ``use'', ``using''. It is worth reminding what the baselines are.}

%The same observations are made for the MFCCs for $k\leq5$. However for a rank $k$ greater than $5$, all the \emph{RbQ} approaches perform equivalently. As such, one can conclude that the use of MFCCs allows us to retrieve closely related scenes, but their lack of discriminative capabilities making the cluster selection/weighting step useless for not so closely related scenes of the same class.

%This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless. \ja{What is the ``at some point'' referring to? In other words, why do we expect this to happen for larger $k$ but not smaller?}

%However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}. The latter even outperforms \emph{ob-a} for $k\geq5$. From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class. This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.

>>>>

Results for the acoustic scene similarity retrieval task demonstrate the superiority of the logarithmically compressed scattering features compared to MFCCs.
Combining these with the \emph{RbQ} cluster model, we obtain significant improvements over traditional BOF and summary statistic measures.

@@

As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied to MFCCs. In contrast, its performance decreases compared to the early approach when applied to scattering features. Extensive analysis of those results is out of the scope of this paper, but this likely results from the dimension reduction applied prior to estimating the GMMs in the BOF model.

The early approach being simpler in terms of implementation and runtime complexity, we retain this method as baseline for the remainder of the experiments.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/baselines}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. } \ja{Should we not have $k$ stricly smaller than $10$ here?}}
\label{fig:ASS_-1}
\end{center}
\end{figure}

@@

Figure \ref{fig:ASS_0} shows how logarithmic compression is beneficial irrespective of the similarity model used. The gain is more pronounced when considering alternative similarities. \ja{What's ``alternative'' similarities?}

@@

For the scattering transform, both \emph{RbQ-c} and \emph{RbQ-w} outperform \emph{early}, thus confirming the benefits of using an relevance-based quantization (\emph{RbQ}) to improve the similarity measures between the scenes. However, it is worth noting that \emph{RbQ-a} performs comparably to or worse than \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids. This result is in line with the findings of \cite{lagrange:hal-01082501}. To take advantage of such a representation, we need to select certain representative centroids when comparing quantized objects.

Furthermore, it appears that \emph{RbQ-c} is better able to characterize the classes compared to \emph{RbQ-w}. Although not the only way of incorporating the number of frames associated to each centroid, the earth mover's distance is a rather natural of doing so. Its worse performance therefore suggests that including this information may not always be desirable. Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames. On the contrary, two similar environments may share a lot of similar sound sources with only a few sources discriminating between them.

With $p@5$ as our metric (cf. \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501}), we see that replacing MFCCs by the logarithmically compressed scattering transform increases performance from $0.31$ to $0.49$. In addition, the relevance-based quantization using the closest similarity (\emph{RbQ-c}) further improves the performance to $0.54$ for a global increase of $0.23$. \vl{It is worth reminding what the baselines are.}
